### 2019-12-11 1 - FoSSaCS 2020 - Submitted/Reviews/Response.txt
UDL

First, we would like to thank the three anonymous reviewers for the many useful and interesting observations. We agree with most of them. Below, we only comment on the main points raised by the three reviewers.

Review 2:

*** (1) The authors should compare more in detail their contribution with the existing literature (the shortcut "This is deeply different from what we do here" [p.3] is too coarse, in fact I disagree on this point). The idea of implementing probabilistic lambda-calculi by a deterministic rewriting enjoying a confluence property is already present in the literature, e.g. Faggian and Ronchi della Rocca's [12], but also (not cited) Alberti's PhD thesis "On operational properties of quantitative extensions of λ-calculus" and (also not cited)

We are aware of the work by Faggian and Ronchi della Rocca. The main difference with this work is that the bang operator $!$ in their calculus $\Lambda_\oplus^!$ plays both the role of a marker for duplicability and of a checkpoint for any probabilistic choice "flowing out" of the term (i.e. being fired). In our calculus, we do not control duplication, but we definitely make use of checkpoints, in the form of the $[a]$ operator. Saying it another way, Faggian and Ronchi's work is inspired by lineare logic, while our approach is inspired by deep inference, even though this is, on purpose, not evident in the design of our calculus. We will add a paragraph to compare our work with the existing literature, in particular with Faggian and Ronchi [12]: we can move some proofs in the appendix to make room

*** Leventis' MSCS 2019 paper "A deterministic rewrite system for the probabilistic lambda-calculus"

We are grateful for the reference, which we knew but whose publication we were not aware. Indeed, Leventis gives an account of standard CBN probabilistic lambda-calculus by similar rewriting to ours - in fact, our system simulates Leventis' $M\oplus N$ as $[a].M\oplus_a B. The difference is our decomposition: Leventis can only express CBN reduction, and not CBV as we do, because the box $[a]$ is always next to the probabilistic choice operator $\oplus_a$.

*** (2) The second issue is more problematic from my point of view. The authors claim in the abstract that their calculus "interprets the call-by-name and call-by-value strategies", but, as far as I understand, the paper misses a clear account of these interpretations.

Call-by-name can be interpreted as above. Call-by-value (or, better, an eager normalization strategy) can be captured by making each $\oplus_a$ to occur deeper than the corresponding binder $[a]$. This way, the former can be copied, but each one of these copies solved by referring to *the same* probabilistic event, since they are all bound by $[a]$.

*** Section 7 is devoted to a translation of cbv into their calculus, but I feel this one-page section sketched and much less accurate than the rest of the paper.

We agree: we would have liked to be more precise, but were struggling against time. We have since improved the exposition by using explicit streams in {0,1}^* to represent probabilistic events.

*** Proposition 35 claims a kind of of simulation but the authors do not discuss the converse, neither the asymptotic behaviour of the semantics, even not with examples.

Indeed, there is _almost_ a bisimulation: probabilistic steps in N and projective steps in [[N]]_cbv are in 1-1 correspondence. But [[N]]_cbv allows beta-steps that duplicate a labelled choice, which are unavailable in N. Still, confluence ensures that a normal form of [[N]]_cbv is the translation of the CBV normal form of N. We would be happy to include a further example.

*** I do not see a similar discussion for the lazy cbn or the head-reduction.

[I don't think we can say too much here, but I am not sure. Maybe we can make $[a]$ to be blocked by $\lambda$s, this way modeling lazyness?].
[WBH: We could offer to say something about head reduction, which (I think) we get immediately by using projective and beta-reduction only in our head context H[]. For CBNeed, we could remark that in practice (meaning: Haskell), monads are used precisely because effects such as probabilities are otherwise too unpredictable. Our syntax is, effectively, the same as monadic computation, which we can make precise.]
[Giulio: I agree with Willem, we can say that head reduction can be simulated by closing beta and projective reductions by head contexts, it's immediate. 
We can say something about laziness, but only in a informal way, as future work, because in my opinoin it's risky to be precise at the moment: laziness it's a subtle matter where it's easy to say something wrong.]

Review 3:

*** I was missing a deeper comparison of the merits of the proposed approach with potential alternative solutions based on linear logic or CBPV. Could the latter lead to a simpler reduction system, e.g. without so many instances of copying present in ->_p?

We point to projective reduction as the simpler reduction system.
[Giulio: In my opinion, this answer is too short and hasty. I would point to the answer we give to point 1 of Referee 2, and I would say that our approach is finer than Faggian and Ronchi Della Rocca because we show that we don't need a marker for duplicability to restore confluence in a probabilistic setting.]