### 2025-12-17 4 - AST/FMC-AST.tex
Initial commit forAST

\documentclass{article}

\usepackage[T1]{fontenc}
\usepackage{palatino}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{stmaryrd}
\usepackage{proof}
\usepackage{tikz}
	\usetikzlibrary{calc}

\usepackage{FMC}

% - - - - - - - - - - - - - - -

\theoremstyle{plain}
\newtheorem{theorem}             {Theorem}
\newtheorem{lemma}      [theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition} [theorem]{Definition}

% - - - - - - - - - - - - - - -

%====================================================================================================
\begin{document}

\title{Simple Types for Almost-Sure Termination}

\maketitle

\begin{abstract}
\end{abstract}

%---------------------------------------------------------------------------------------------------- INTRODUCTION

\section{Introduction}

* Conservative extension of [Curry \&\ Howard meet Borel] and [Simple Types for Probabilistic Termination].

* Uses \emph{iteration} rather than \emph{recursion} to obtain probabilistic typeability.

* Probabilistic recursion is difficult as it requires probabilistic function variables at type level, essentially using dependent types. Example: probabilistic Church numerals and Y-combinator.

* Probabilistic iteration is straightforward and natural, taking  $f : A \to A+B$ to $\mathsf{iter}~f : A\to B$, which is AST if $A+B$ is a probabilistic sum which has a non-zero probability of choosing $B$.

%---------------------------------------------------------------------------------------------------- PROBABILISTIC FMC

\section{The probabilistically-typed FMC}

Terms:
\[
\term{M,N} ~\coloneqq~ \term{x} ~\mid~ \term{<x>.M} ~\mid~ \term{[N].M} ~\mid~ \term{i} ~\mid~ \term{M;i->N} ~\mid~ \term{M^i} ~\mid~ \term{?a.M}
\]

\noindent
Abbreviations:
\[
	\term{M+N} ~=~\term{?a.a;\top->M;\bot->N}
\]

\noindent
Sets of things:
\[
\begin{array}{lll}
	\0,\1   &~\in~\B      & \text{boolean constants}
\\	a,x,y,z &~\in~\Var    & \text{variables}
\\  i,j,k   &~\in~\Choice & \text{choices, where also}~\B\subset\Choice
\\	R,S,T   &~\in~\Stack  & \text{stacks of terms}
\\  L,K     &~\in~\Cont   & \text{continuation stacks: stacks of cases}~\term{i->M}
\\	r,s,t   &~\in~\Trace  & \text{traces: finite sequences of booleans}~\top,\bot
\\  p,q     &~\in~[0,1]   & \text{probabilities}
\end{array}
\]

Machine:
\[
\begin{array}{lc}
	  \mathsf{push}     & \step- t S {[N].M} K             t {S\,\term N} M K
\\ \\ \mathsf{pop}      & \step- t {S\,\term N} {<x>.M} K  t S {\{N/x\}M} K
\\ \\ \mathsf{select}   & \step- t S i {(\term{i->M})\,K}  t S M K
\\ \\ \mathsf{reject}   & \step- t S i {(\term{j->M})\,K}  t S i K
\\ \\ \mathsf{sequence} & \step- t S {M;i->N} K            t S M {(\term{i->N})\,K}
\\ \\ \mathsf{iterate}  & \step- t S {M^i} K               t S M {(\term{i->M^i})\,K}
\\ \\ \mathsf{sample}   & \step- t S {?a.M} K        {t\,\1} S {\{\1/a\}M} K
                   \quad  \step- t S {?a.M} K        {t\,\0} S {\{\0/a\}M} K
\end{array}
\]

A \emph{run} of the machine is a sequence of steps, written with a double line as below. A \emph{successful} run is one terminating in a \emph{final state} $\state tTi\e$ with a choice term $\term i$ and empty continuation stack.
\[ 
	\step= sSMK tTNL
\]

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - Big-step semantics
\subsection{Big-step semantics}

A \emph{frontier} $F$ is a set of traces such that no element in $F$ is a prefix of another. A frontier is \emph{total} if it contains a prefix of every stream in $\B^\N$.

To collect the outcomes of evaluating a term, a \emph{return family} $\DS$ in a frontier $F$ assigns to each trace $t\in F$ a pair $(S,\term i)$ of a stack and choice. We write $\floor\DS$ for the frontier $F$ over which $\DS$ is defined.
\[
	\DS,\DT~:~F\to\Stack\times\Choice
\]
The set $\runs SM$ for a stack $S$ and term $\term M$ is the return family that collects the traces and outputs of all successful runs, as follows. Note that this is well-defined as a family over a frontier since each run has a unique trace.
\[
	\runs SM~=~\{~t\mapsto (T,\term i)~\mid~\step= \e SM\e tTi\e~\}
\]

\emph{Union} of two return families $\DS\cup\DT$ is that of their underlying relations, and is a return family if they are defined over disjoint frontiers, $\floor\DS\cap\floor\DT=\varnothing$. Likewise, \emph{inclusion} between families $\DS\subseteq\DT$ is inclusion as relations, which means $\floor\DS\subseteq\floor\DT$ and $\DS(t)=\DT(t)$ for all $t\in\floor\DS$. We define the following notions below: the \emph{extension} $s\pfx\DS$ of a return family $\DS$ with a trace $s$, its \emph{restriction} $\DS\only i$ to a choice $i$, and its \emph{exclusion} $\DS\less i$ of $i$. 
\[
\begin{array}{l@{}l}
	    s\pfx\DS   &~=~\{ s\,t\mapsto(S,j) \mid t\mapsto(S,j)\in\DS\}
\\[5pt]	\DS\only i &~=~\{ t\mapsto(S,j)\in\DS \mid j=i \} 
\\[5pt]	\DS\less i &~=~\{ t\mapsto(S,j)\in\DS \mid j\neq i \}
\end{array}
\]

\begin{definition}
The \emph{evaluation} relation $\evalarrow$ on terms $\eval SM\DT$ and continuations $\eval\DS{i->M}\DT$ is defined inductively by the following rules.
\[
\begin{array}{@{}c@{\qquad}c@{\qquad}c@{}}
	\infer{\eval Si{\{\e\mapsto (S,i)\}}}{}
&	\infer{\eval {S\,\term N}{<x>.M}\DT}{\eval S{\{N/x\}M}\DT}
&	\infer{\eval R{M;i->N}\DT}{\eval RM\DS && \eval\DS{i->N}\DT}
\\ \\
	\infer{\eval SM\varnothing}{}
&	\infer{\eval S{[N].M}\DT}{\eval {S\,\term N}M\DT}
&	\infer{\eval S{?a.M}{(\0\pfx\DT_0)\cup(\1\pfx\DT_1)}}{\eval S{\{\0/a\}M}{\DT_0} && \eval S{\{\1/a\}M}{\DT_1}}
\\ \\
\multicolumn{2}{@{}c@{\qquad}}{
	\infer{\eval S{M^i}{\sup_{n\in\N}(\DT_n\less i)}}{
    	\eval SM{\DT_0}
    	&& 
    	\{\eval{\DT_n}{i->M}{\DT_{n+1}}\}_{n\in\N}
    }
}&
	\infer{\eval\DS{i->M}{\DS\less i\cup\bigcup_{t\in\floor{\DS\only i}} t\pfx\DT_t}}{ \{\eval {S_t}M{\DT_t}\}_{t\,\mapsto\,(S_t,i)~\in~\DS\only i} } 
\end{array}	
\]
\end{definition}

Observe that the \emph{sample} and \emph{sequence} rules are well-defined as the unions are disjoint, and in the \emph{iterate} rule for $\term{M^i}$ the supremum is well-defined since each $\DT_{n+1}$ is given (in the continuation rule) as $\DT_n\less i\cup\DT'$ for some $\DT'$, so that we immediately have $\DT_n\less i\subseteq\DT_{n+1}\less i$.

% - - - - - - - - - - - - - - -
\begin{proposition}
\label{prop:eval-run}
If $\eval SM\DT$ then $\DT~\subseteq~\runs SM$.
\end{proposition}

\begin{proof}
By induction on the derivation for $\eval SM\DT$.
\begin{itemize}

	% Choice
	\item \emph{Choice:}
\[
	\infer{\eval Si{\D{\e\mapsto (S,i)}}}{}
\]
The only run for the state $(\e,S,\term i,\e)$ is the empty run, which gives the singleton $\runs SM=\{\e\mapsto (S,i)\}$ as required.

	% Fail
	\item \emph{Fail:}
\[
	\infer{\eval SM\varnothing}{}
\]
It is immediate that $\varnothing\subseteq\runs SM$.

	% Pop
	\item \emph{Pop:}
\[
	\infer{\eval {S\,\term N}{<x>.M}\DT}{\eval S{\{N/x\}M}\DT}
\]
Since a run for $\state \e{S\,\term N}{<x>.M}\e$ must necessarily start with a $\mathsf{pop}$ transition to $\state\e S{\{N/x\}M}\e$, we have:
\[
	\runs{S\,\term N}{<x>.M}~=~\runs S{\{N/x\}M}
\]
The case is then immediate by induction.

	% Push
	\item \emph{Push:}
\[
	\infer{\eval S{[N].M}\DT}{\eval {S\,\term N}M\DT}
\]
Since a run for $\state \e S{[N].M}\e$ must necessarily start with a $\mathsf{push}$ transition to $\state\e{S\,\term N}M\e$, we have:
\[
	\runs S{[N].M}~=~\runs{S\,\term{N}}M
\]
The case is then immediate by induction.

	% Sequence
	\item \emph{Sequence:}
\[
	\infer{\eval R{M;i->N}\DT}{\eval RM\DS && \eval\DS{i->N}\DT}
\qquad
	\infer{\eval\DS{i->M}{\DS\less i\cup\bigcup_{t\in\floor{\DS\only i}} t\pfx\DT_t}}{ \{\eval {S_t}M{\DT_t}\}_{t\,\mapsto\,(S_t,i)~\in~\DS\only i} }
\]
Let $s\mapsto (T,j)\in\DT$ as above left, where $\DT=\DS\less i\cup\bigcup_{t\in\floor{\DS\only i}} t\pfx\DT_t$ as above right. It must be shown that $s\mapsto (T,j)\in\runs R{M;i->N}$. There are two cases.

\begin{itemize}
	\item $s\mapsto(T,j)\in\DS\less i$. Then $j\neq i$ and by induction on $\eval RM\DS$ we have $s\mapsto(T,j)\in\runs RM$, which means there is a run as below left. This gives the run below right, which shows $s\mapsto(T,j)\in\runs R{M;i->N}$.
\[
	\step= \e RM\e tTj\e
\qquad\qquad	    	
    \begin{array}{@{(~}l@{~,~}l@{~,~}r@{~,~}r@{~)}}
    	      \e & R & \term{M;i->N} & \e
    \\ \hline \e & R & \term{M}      & (\term{i->N})
    \\ \dline  s & T & \term{j}      & (\term{i->N})
    \\ \hline  s & T & \term{j}      & \e
    \end{array}
\]

	\item $s\mapsto(T,j)\in\bigcup_{t\in\floor{\DS\only i}} t\pfx\DT_t$. Then there is some $t\mapsto(S,i)\in\DS\only i$ such that $s = t\,t'$ and $t'\mapsto(T,j)\in\DT_t$. By induction, $t\mapsto(S,i)\in\runs RM$ and $t'\mapsto(T,j)\in\runs SN$. This gives the two runs below left, from which that below right is then constructed, showing that $s\mapsto(T,j)\in\runs R{M;i->N}$.
\[
	\begin{array}{@{}c@{}}
		\step= \e RM\e tSi\e
	\\ \\
		\step= \e SN\e{t'}Tj\e
	\end{array}
\qquad\qquad
    \begin{array}{@{(~}l@{~,~}l@{~,~}r@{~,~}r@{~)}}
    	      \e    & R & \term{M;i->N} & \e
    \\ \hline \e    & R & \term{M}      & (\term{i->N})
    \\ \dline t     & S & \term{i}      & (\term{i->N})
    \\ \hline t     & S & \term{N}      & \e
    \\ \dline t\,t' & T & \term{j}      & \e
    \end{array}
\]
\end{itemize}

	% Sample
	\item \emph{Sample:}
\[
	\infer{\eval S{?a.M}{(\0\pfx\DT_0)\cup(\1\pfx\DT_1)}}{\eval S{\{\0/a\}M}{\DT_0} && \eval S{\{\1/a\}M}{\DT_1}}
\]
Let $t\mapsto(T,i)\in\0\pfx\DT_0\cup\1\pfx\DT_1$. Then either $t=\0\,s$ and $s\mapsto(T,i)\in\DT_0$, for some $s$, or $t=\1\,s$ and $s\mapsto(T,i)\in\DT_1$. We consider the case $t=\0\,s$; the other is analogous. By induction we have the run below left, which then gives that below right, showing that $t\mapsto(T,i)\in\runs S{?a.M}$.
\[
	\step= \e S{\{\0/a\}M}\e sTi\e 
\qquad\qquad
    \begin{array}{@{(~}l@{~,~}l@{~,~}r@{~,~}r@{~)}}
    	      \e    & S & \term{?a.M}      & \e
    \\ \hline \0    & S & \term{\{\0/a\}M} & \e
    \\ \dline \0\,s & T & \term{i}         & \e
    \end{array}
\]

	% Iterate
	\item \emph{Iterate:}
\[
	\infer{\eval S{M^i}{\sup_{n\in\N}(\DT_n\less i)}}{
    	\eval SM{\DT_0}
    	&& 
    	\{\eval{\DT_n}{i->M}{\DT_{n+1}}\}_{n\in\N}
    }
\qquad
	\infer{\eval\DS{i->M}{\DS\less i\cup\bigcup_{t\in\floor{\DS\only i}} t\pfx\DT_t}}{ \{\eval {S_t}M{\DT_t}\}_{t\,\mapsto\,(S_t,i)~\in~\DS\only i} } 
\]
First, by induction on $n$ it will be shown that if $t\mapsto(T,j)\in\DT_n$ then there is a run:
\[
	\step= \e S{M^i}\e tTj{\term{i -> M^i}} 
\]
For $n=0$, by the inductive hypothesis for $\eval SM{\DT_0}$ and $t\mapsto(T,j)\in\DT_0$ we have the run below left, which then gives that below right.
\[
	\step= \e SM\e tTj\e 
\qquad	
    \begin{array}{@{(~}l@{~,~}l@{~,~}r@{~,~}r@{~)}}
    	      \e & S & \term{M^i} & \e
    \\ \hline \e & S & \term{M}   & (\term{i->M^i})
    \\ \dline  t & T & \term{j}   & (\term{i->M^i})
    \end{array}
\]
For the case $n+1$, by the inductive hypothesis for $\eval{\DT_n}{i->M}{\DT_{n+1}}$ and $t\mapsto(T,j)\in\DT_{n+1}$ we have two cases: $t\mapsto(T,j)$ is either in $\DT_n\less i$ or in $\bigcup_{t\in\floor{\DT_n\only i}} t\pfx\DT_t$. The former case is immediate by the inductive hypothesis. In the latter case, $t=r\,s$ and there are $R$ and $\DT_s$ such that
\[
	r\mapsto(R,i)\in\DT_n
\qquad
	\eval RM{\DT_s}
\qquad
	s\mapsto(T,j)\in\DT_s
\]
The first statement gives the first run below left, by the induction on $n$, and the induction on the latter two statements gives the second. Then the required run is given below right.
\[
	\begin{array}{@{}c@{}}
		\step= \e S{M^i}\e rRi{(\term{i->M^i})}
	\\ \\
		\step= \e RM\e sTj\e
	\end{array}
\qquad\qquad
    \begin{array}{@{(~}l@{~,~}l@{~,~}r@{~,~}r@{~)}}
    	      \e    & S & \term{M^i} & \e
    \\ \dline r     & R & \term{i}   & (\term{i->M^i})
    \\ \hline r     & R & \term{M^i} & \e
    \\ \hline r     & R & \term{M}   & (\term{i->M^i})
    \\ \dline r\,s  & T & \term{j}   & (\term{i->M^i})
    \end{array}
\]

For the overall case, let $t\mapsto(T,j)\in\sup_{n\in\N}(\DT_n\less i)$. Then there is some $m$ such that $t\mapsto(T,j)$ is in $\DT_m\less i$. By the above, and since $i\neq j$, we have the following run, which shows that $t\mapsto(T,j)\in\runs S{M^i}$.
\[
    \begin{array}{@{(~}l@{~,~}l@{~,~}r@{~,~}r@{~)}}
    	      \e & S & \term{M^i} & \e
    \\ \dline  t & T & \term{j}   & (\term{i->M^i})
    \\ \hline  t & T & \term{j}   & \e
    \end{array}
\]

\end{itemize}
\end{proof}
% - - - - - - - - - - - - - - -

{\color{green!66!black} The following is not essential but could be nice. Seems a bit of work, though.}

\begin{proposition}
$\eval SM{\runs SM}$.
\end{proposition}

%---------------------------------------------------------------------------------------------------- TYPES

\section{Types}

Types will use \emph{indexed probability (sub-)distributions}, which are families over finite sets of choices $I,J$.

\noindent
Types:
\[
\begin{array}{lr@{~\coloneqq~}l}
	\text{types}	   & \type{s,t} & \type{!s => !tI}
\\	\text{stack types} & \type{!t}  & \type{t_1..t_n}
\\  \text{sum types}   & \type{!tI} & \type{\{!t_i.p_i\}_{i\in I}}~\quad~\sum_{i\in I}p_i\leq 1
\end{array}
\]

\noindent
Definitions:
\[
\begin{array}{l@{}ll}
		\type{(!t.p)_i}    &~=~ \type{\{!t_i.p_i\}_{\{i\}}}                       & \text{where~}\type{!t_i}=\type{!t},~p_i=p
\\[5pt] \type{!sI + !tJ}   &~=~ \type{!sI \cup !tJ}                               & \text{if~}(I\cap J=\varnothing)
\\[5pt] \type{!s\,!tI}     &~=~ \type{\{!s\,!t_i . p_i \mid !t_i.p_i \in !tI\}}
\\[5pt]	\type{q.!tI}       &~=~ \type{\{ !t_i . q\,p_i \mid  !t_i.p_i \in !tI\}}
\\[5pt]	\type{!tI-J}       &~=~ \type{\{ !t_i.p_i \in !tI \mid i\notin `J\}}
\\[5pt] \type{!tI|J}       &~=~ \type{\{ !t_i.p_i \in !tI \mid i\in `J\}}
\\[5pt] \type{!sI ++ !tJ}  & \multicolumn{2}{@{}l@{}}{~=~ \type{\{!t_i . p_i+q_i \mid !t_i.p_i \in !sI|J~`,~!t_i.q_i\in!tJ|I\}}}
\\[5pt]                    &\quad\type{\cup~(!sI-J)~\cup~(!tJ-I)}                 & \text{if~} \forall i\in(I\cap J).~\type{s_i}=\type{t_i}
\end{array}
\]

%\[
%\begin{array}{lr@{~\coloneqq~}l}
%	\text{types}	   & \type{s,t} & \type{!s => !tI}
%\\	\text{stack types} & \type{!t}  & \type{t_1..t_n}
%\\  \text{sum types}   & \type{!tI} & \type{\{i \mapsto !t_i.p_i\}_{i\in I}}~\quad~\sum_{i\in I}p_i\leq 1
%\end{array}
%\]
%
%\noindent
%Definitions:
%\[
%\begin{array}{lll}
%		\type{i(!t.p)}     &~=~ \type{\{i \mapsto !t.p\}}
%\\[5pt] \type{!sI ++ !tJ}  &~=~ \type{!sI \cup !tJ} \qquad (I\cap J=\varnothing)
%\\[5pt] \type{!s\,!tI}     &~=~ \type{\{i\mapsto !s\,!t_i . p_i \mid i\mapsto !t_i.p_i \in !tI\}}
%\\[5pt]	\type{q.!tI}       &~=~ \type{\{i\mapsto  !t_i . q\,p_i \mid i\mapsto !t_i.p_i \in !tI\}}
%\\[5pt]	\type{!tI-J}       &~=~ \type{\{i\mapsto !t_i.p_i \in !tI \mid i\notin J\}}
%\\[5pt] \type{!tI|J}       &~=~ \type{\{i\mapsto !t_i.p_i \in !tI \mid i\in J\}}
%\\[5pt] \floor{\type{!tI}} &~=~ \{ i\mapsto \type{!t_i}\}_{i\in I}
%\\[5pt] \type{!sI + !tJ}   &~=~ \type{\{i\mapsto !t_i . p_i+q_i \mid i\mapsto !t_i.p_i \in !sI|J~`,~i\mapsto !t_i.q_i\in!tJ|I\}}
%\\                         &\qquad \type{~\cup~(!sI-J)~\cup~(!tJ-I)} 
%%\\[5pt] \type{!sI + !tI}   &~=~ \type{\{i\mapsto !t_i.p_i+q_i \mid i\mapsto !t_i.p_i \in !sI,~i\mapsto !t_i.q_i\in !tI \}} & \floor{\type{!sI}}=\floor{\type{!tI}}
%%\\[5pt] \type{!sI + !tJ}   &~=~ \type{(!sI|J v !tJ|I) u (!sI-J) u (!tJ-I)}
%\end{array}
%\]

\noindent
Typing rules:
\[
\begin{array}{lc}
      \mathsf{variable}  & \infer{\term{G, x:t |- x:t}}{}
\\ \\ \mathsf{expansion} & \infer{\term{G |- M: !s\,!r => !r\,!tI}}{\term{G |- M: !s => !tI}}
\\ \\ \mathsf{push}      & \infer{\term{G |- [N].M : !s => !tI}}{\term{G |- N:r} && \term{G |- M : r\,!s => !tI}}
\\ \\ \mathsf{pop}       & \infer{\term{G |- <x>.M : r\,!s => !tI}}{\term{G , x:r |- M: !s => !tI}}
\\ \\ \mathsf{choice}    & \infer{\term{G |- i: !s => (!s.1)_i}}{}
\\ \\ \mathsf{sequence}  & \infer{\term{G |- M;i->N : !r => !sI ++ (p . !tJ)}}
                                 {\term{G |- M: !r => !sI + (!s.p)_i} && 
                                  \term{G |- N: !s => !tJ}}
\\ \\ \mathsf{iterate}   & \infer{\term{G |- M^i: !s => /1{1-p} . !tI}}{\term{G |- M: !s => !tI + (!s.p)_i} && (p\neq 1)}
\\ \\ \mathsf{sample}    & \infer{\term{G |- ?a.M : !r => (/12 . !sI) ++ (/12 . !tJ)}}{\term{G , a: \e => (\e.1)_\bot |- M: !r => !sI} && \term{G , a: \e => (\e.1)_\top |- M: !r => !tJ}}
\\ \\ \mathsf{zero}      & \infer{\term{G |- M : !s => 0 . !tI}}{}
\end{array}
\]

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - Runnability
\newpage
\subsection{Runnability}

The \emph{probability} of a trace $t$ is $p(t)=(\fraction12)^{|t|}$ where $|t|$ is the \emph{length} of $t$.

\[
\begin{array}{l@{}l}
%	     |\DS|       &~=~ \sum_{t\in\floor\DS}~p(t)
%\\[5pt]
         |\DS|_i     &~=~ \sum_{t\in\floor{\DS\only i}}~p(t)
\\[5pt]  \floor\DS_i &~=~ \{ S \mid t\mapsto(S,i)\in\DS \}
\end{array}
\]

\begin{definition}
The set $\RUN{t}$ of \emph{runnable terms} for a type $\type t$ is defined as the set of closed terms
\[
	\RUN{!s => !tI} = \{ \term M ~\mid~ \forall S \in\RUN{!s}.~\exists\DT\in\RUN{!tI}.~\eval SM\DT~\}
\]
where the runnable sets for stack types $\type{!t}$ and sum types $\type{!tI}$ are as follows.
\[
    \RUN{t_1..t_n} = \{ \e\,\term{M_1}\dots \term{M_n} \mid \term{M_i}\in\RUN{t_i} \}
\]
\[
	\RUN{\{!t_i.p_i\}_{i\in I}} = \{ \DT \mid \forall i\in I.~\floor\DT_i\subseteq\RUN{!t_i}~\wedge~|\DT|_i\geq p_i~\}
\]
\end{definition}

% - - - - - - - - - - - - - - - 
\begin{lemma}
\label{lem:run-union}
If $\DS\in\RUN{!sI}$ and $\DT\in\RUN{!tJ}$ then $\DS\cup\DT\in\RUN{!sI ++ !tJ}$ if $\DS\cup\DT$ and $\type{!sI++!tJ}$ are defined.
\end{lemma}

\begin{proof}
Let $\type{!sI}=\type{\{!s_i.p_i\}_{i\in I}}$ and $\type{!tJ}=\type{\{!t_j.q_j\}_{j\in J}}$. Note that $\DS\cup\DT$ is defined if $\floor\DS\cap\floor\DT=\varnothing$, and $\type{!sI++!tJ}$ if $\type{!s_i}=\type{!t_i}$ for all $i\in I\cap J$. We need to show two properties.
\begin{itemize}
	 \item[1)] If $t\mapsto(S,i)\in\DS\cup\DT$ then $S\in\RUN{!s_i}$ if $i\in I$ and $S\in\RUN{!t_i}$ if $i\in J$. This is as follows: since $\DS\in\RUN{!sI}$, if $t\mapsto(S,i)\in\DS$ then $S\in\RUN{!s_i}$, and since $\DT\in\RUN{!tJ}$, if $t\mapsto(S,i)\in\DT$ then $T\in\RUN{!t_j}$.

	 \item[2)] For $i\in I\cup J$, the probability $|\DS\cup\DT|_i$ is at least: $p_i+q_i$ if $i\in I\cap J$, $p_i$ if $i\in I\less J$, and $q_i$ if $i\in J\less I$. This is as follows: first, $|\DS\cup\DT|_i=|\DS|_i+|\DT|_i$, since both are the sum over those $p(t)$ such that $t\in\floor{\DS\only i}$ or $t\in\floor{\DT\only i}$. Then since $\DS\in\RUN{!sI}$ the probability $|\DS|_i\geq p_i$ for $i\in I$, and since $\DT\in\RUN{!tJ}$ the probability $|\DT|_i\geq q_i$ for $i\in J$.
\end{itemize}
\end{proof}
% - - - - - - - - - - - - - - -

For a context $\Gamma=\term{x_1:t_1,..,x_n:t_n}$ the set $\SRUN{G}$ is a set of substitution maps $\mu$ that assigns each $x_i$ a runnable term $\term M\in\RUN{t_i}$.
\[
	\SRUN{G}~=~\{ \mu \mid \forall\term{x:t}\in\Gamma.~\term{\mu x}\in\RUN{t}\}
\]

% - - - - - - - - - - - - - - -
\begin{lemma}
\label{lem:run}
If $\term{G |- M:t}$ and $\mu\in\SRUN{G}$ then $\term{\mu M}\in\RUN{t}$.
\end{lemma}

\begin{proof}

By induction on the typing derivation.
\begin{itemize}

	% Variable
	\item \emph{Variable:} 
\[
	\infer{\term{G, x:t |- x:t}}{}
\]
By the definition of $\SRUN{G}$ we have $\term{\mu x}\in\RUN t$.

	% Choice
	\item \emph{Choice:}
\[
\infer{\term{G |- i: !s => (!s.1)_i}}{}
\]
Given $S\in\RUN{!s}$ we have $\eval Si\DT$ where $\DT={\{\e\mapsto(S,i)\}}$. Then $\floor\DT_i=\{S\}\subseteq\RUN{!s}$ and $|\DT|_i=p(\e)=1$, which gives $\DT\in\RUN{(!s.1)_i}$, and hence $\term{\mu i}=\term i\in\RUN{!s=>(!s.1)_i}$.

	% Zero
	\item \emph{zero}
\[
\infer{\term{G |- M : !s => 0 . !tI}}{}
\]
We have $\eval S{\mu M}\varnothing$ for any $S$ and $\mu$, and $\varnothing\in\RUN{0.!tI}$ since $\floor\varnothing_i=\varnothing\subseteq\RUN{!t_i}$ and $|\varnothing|_i\geq 0$ for all $i$.

	% Expansion
	\item \emph{Expansion:}
\[
\infer{\term{G |- M: !s\,!r => !r\,!tI}}{\term{G |- M: !s => !tI}}
\]
Let $R\in\RUN{!r}$, $S\in\RUN{!s}$, and $\mu\in\SRUN{G}$. By induction $\eval S{\mu M}\DT$ for some $\DT\in\RUN{!tI}$. Then by induction on the evaluation relation, $\eval{RS}{\mu M}{R\DT}$ where $R\DT=\{ t \mapsto (RT,i) \mid t\mapsto (T,i)\in\DT\}$. Since $R\in\RUN{!r}$ it follows by the definition of $\RUN{!tI}$ that $R\DT\in\RUN{!r\,!tI}$.

	% Push
	\item \emph{Push:} 
\[
\infer{\term{G |- [N].M : !s => !tI}}{\term{G |- N:r} && \term{G |- M : r\,!s => !tI}}
\]
Let $S\in\RUN{!s}$ and $\mu\in\SRUN{G}$. By induction $\term{\mu N}\in\RUN{r}$ and $\term{\mu M}\in\RUN{r\,!s=>!tI}$, where by definition we get $\eval{S\,\term{\mu N}}{\mu M}\DT$ for some $\DT\in\RUN{!tI}$. By the definition of evaluation we get $\eval{S}{[\mu N].\mu M}\DT$ so that we have $\term{[\mu N].\mu M}=\term{\mu([N].M)}\in\RUN{!s=>!tI}$.

	% Pop
	\item \emph{Pop:}
\[
\infer{\term{G |- <x>.M : r\,!s => !tI}}{\term{G , x:r |- M: !s => !tI}}
\]
Let $S\in\RUN{!s}$, $\term N\in\RUN r$, and $\mu\in\SRUN{G}$. Let $\mu'=\mu\{\term N/x\}$, the substitution map that sends $x$ to $\term N$ and every other variable $y$ to $\mu y$, so that $\mu'\in\SRUN{G,x:r}$. By induction $\eval S{\mu'M}\DT$ for some $\DT\in\RUN{!tI}$, and since $\term{\mu'M}=\term{\mu(\{N/x\}M)}$ by the definition of evaluation $\eval{S\,\term N}{\mu(<x>.M)}\DT$. It follows that $\term{\mu(<x>.M)}\in\RUN{r\,!s=>!tI}$.

	% Sample
	\item \emph{Sample:}
\[
\infer{\term{G |- ?a.M : !r => (/12 . !sI) ++ (/12 . !tJ)}}{\term{G , a: \e => (\e.1)_\0 |- M: !r => !sI} && \term{G , a: \e => (\e.1)_\1 |- M: !r => !tJ}}
\]
Let $R\in\RUN{R}$ and $\mu\in\SRUN{G}$. Let $\mu_0=\mu\{\term\0/a\}$ and $\mu_1=\mu\{\term\1/a\}$, so that $\mu_0\in\SRUN{G , a: \e => (\e.1)_\0}$ and $\mu_1\in\SRUN{G , a: \e => (\e.1)_\1}$. By induction $\eval R{\mu_0M}{\DT_0}$ and $\eval R{\mu_1M}{\DT_1}$ where $\DT_0\in\RUN{!sI}$ and $\DT_1\in\RUN{!tJ}$. We have the following evaluation.
\[
\infer{\eval S{\mu(?a.M)}{(\0\pfx\DT_0)\cup(\1\pfx\DT_1)}}{\eval S{\mu\{\0/a\}M}{\DT_0} && \eval S{\mu\{\1/a\}M}{\DT_1}}
\]
Since $\0\pfx\DT_0$ extends every trace in $\DT_0$ by $\bot$, it multiplies every probability $|\DT_0|_i$ by $\fraction 12$, so that we have $\0\pfx\DT_0\in\RUN{/12.!sI}$. Likewise we have $\1\pfx\DT_1\in\RUN{/12.!tJ}$. Then by Lemma~\ref{lem:run-union}
$(\0\pfx\DT_0)\cup(\1\pfx\DT_1)\in\RUN{(/12 . !sI) ++ (/12 . !tJ)}$ and hence $\term{\mu(?a.M)}\in\RUN{!r => (/12 . !sI) ++ (/12 . !tJ)}$.

	% Continuation
	\item \emph{Continuation:}
As a lemma for the remaining two cases, it will be shown that if $\DS\in\RUN{!sI + (!s.p)_i)}$ and $\term{M}\in\RUN{!s => !tJ}$ then $\eval\DS{i->M}\DT$ for some $\DT\in\RUN{!sI ++ (p . !tJ)}$. By definition, for every $t\mapsto(S_t,i)\in\DS$ we have $S_t\in\RUN{!s}$, and by induction $\eval{S_t}M{\DT_t}$ for some $\DT_t\in\RUN{!tJ}$. Evaluation is then as follows.
\[
\infer{\eval\DS{i->M}{\DS\less i\cup\bigcup_{t\in\floor{\DS\only i}} t\pfx\DT_t}}{ \{\eval {S_t}M{\DT_t}\}_{t\,\mapsto\,(S_t,i)~\in~\DS\only i} } 
\]
Let $\DT=\bigcup_{t\in\floor{\DS\only i}}t\pfx\DT_t$. Since $\DT_t\in\RUN{!tJ}$ and prefixing each trace in $\DT_t$ with $t$ multiplies each probability $|\DT_t|_i$ by $p(t)$, we have $t\pfx\DT_t\in\RUN{p(t).!tJ}$. Then since $|\DS|_i=\sum\{p(t)\mid t\in\floor{\DS\only i}\}\geq p$ we have $\DT\in\RUN{p.!tJ}$. By Lemma~\ref{lem:run-union} then $\DS\less i\cup\DT\in\RUN{!sI ++ (p.!tJ)}$.

	% Sequence
	\item \emph{Sequence:} 
\[
\infer{\term{G |- M;i->N : !r => !sI ++ (p . !tJ)}}
      {\term{G |- M: !r => !sI + (!s.p)_i} && 
       \term{G |- N: !s => !tJ}}
\]
Let $R\in\RUN{!r}$ and $\mu\in\SRUN{G}$. By induction we have $\eval R{\mu M}\DS$ where $\DS\in\RUN{!sI+(!s.p)_i}$. By the \emph{continuation} case above, $\eval\DS{i->\mu N}\DT$ for some $\DT\in\RUN{!sI ++ (p . !tJ)}$. The case then follows by the evaluation below, which gives that $\term{\mu(M;i->N)}\in\RUN{!r => !sI ++ (p . !tJ)}$
\[
	\infer {\eval R{\mu M;i->\mu N}\DT} {\eval R{\mu M}\DS && \eval\DS{i->\mu N}\DT}
\]

	% Iterate
	\item \emph{Iterate:}
\[
\infer{\term{G |- M^i: !s => /1{1-p} . !tI}}{\term{G |- M: !s => !tI + (!s.p)_i} && (p\neq 1)}
\]
Let $S\in\RUN{!s}$ and $\mu\in\SRUN{G}$ so that by induction $\eval S{\mu M}{\DT_0}$ for some $\DT_0\in\RUN{!tI + (!s.p)_i}$. By induction on $n$, using the \emph{continuation} case above, there are $\DT_n$ for all $n\in\N$ where $\eval{\DT_n}{i->\mu M}{\DT_{n+1}}$, with the following evaluation.
\[
\infer{\eval S{\mu M^i}{\sup_{n\in\N}(\DT_n\less i)}}{
    	\eval S{\mu M}{\DT_0}
    	&& 
    	\{\eval{\DT_n}{i->\mu M}{\DT_{n+1}}\}_{n\in\N}
    }
\]
Let $p_n=\sum_{m\leq n}p^m$. By induction on $n$ we have 
\[
	\DT_n\in\RUN{(p_n\cdot !tI) + p^n\cdot (!s.p)_i)}
\] 
as follows. For $n=0$ we have $\DT_0\in\RUN{!tI + (!s.p)_i}$, and for $n+1$ we have
\[
	\type{(p_n . !tI) + p^n . (p . (!tI + (!s.p)_i)))}
	~=~
	\type{(p_{n+1} . !tI)+p^{n+1}.(!s.p_i)} 
\]
so that by the \emph{continuation} case $\DT_{n+1}\in\RUN{(p_{n+1} . !tI)+p^{n+1}.(!s.p_i)}$. Since $i\notin I$ we have $\DT_n\less i\in\RUN{p_n.!tI}$, and since $\{p_n\mid n\in\N\}$ is a geometric series with limit $\fraction 1{1-p}$ we have that $\sup_{n\in\N}(\DT_n\less i)\in\RUN{/1{1-p}.!tI}$ and $\term{\mu M^i}\in\RUN{/1{1-p}!tI}$.
\end{itemize}
\end{proof}
% - - - - - - - - - - - - - - -

Let $|\type{!tI}|=\sum\{ p_i \mid \type{!t_i.p_i} \in \type{!tI} \}$ and $|\DT|=\sum\{ p(t)\mid t\in\floor\DT\}$.

% - - - - - - - - - - - - - - -
\begin{theorem}[Almost-sure termination]
If $\term{|- M: \e => !tI}$ then $\term M$ almost-surely terminates with probability $|\type{!tI}|$.
\end{theorem}

\begin{proof}
Let $\type{!tI}=\type{\{!t_i.p_i\}_{i\in I}}$. By Lemma~\ref{lem:run} we have $\eval\e M\DT$ with $\DT\in\RUN{!tI}$, which means that $|\DT|_i\geq p_i$ for every $i\in I$, so that $|\DT|\geq|\type{!tI}|$. By Proposition~\ref{prop:eval-run} $\DT\subseteq\runs\e M$, so that $|\runs\e M|\geq|\type{!tI}|$.
\end{proof}
% - - - - - - - - - - - - - - -

%----------------------------------------------------------------------------------------------------
\newpage
\subsection{Encoding Markov chains}

A Markov chain
\[
	C~=~(S,~s\in S,~F\subseteq S,~\delta:S\less F\to\DD{S})
\]
is given by a set of \emph{states} $S$, a \emph{starting state} $s$, a set of \emph{final states} $F$, and a stochastic transition function $\delta$ from non-final states to states.

\newcommand\sem[1]{\llbracket#1\rrbracket}

The interpretation $\sem C$ of a Markov chain as a term is given as follows. For $C$ as above and $t\in S\less\{s\}$, let
\[
	C(t)~=~(S,~t,~F\cup\{s\},~\delta\less\{s\})
\]
where $\delta\less X$ is the transition function $\delta$ restricted to the domain $\mathrm{dom}(\delta)\less X$. For a distribution of terms
\[
	\DM~=~\D{{\term{M_1}}.p_1,..,{\term{M_n}}.p_n}
\]
let the term $\term{\lfloor\DM\rfloor}$ be the internalization of that distribution:
\[
	\lfloor\D{{\term{M}}.1}\rfloor~=~\term{M} 
\qquad\qquad
	\lfloor\D{{\term{M}}.p}+\DM\rfloor~=~\term{?a_p.a ; \top-> M; \bot -> \lfloor{\fraction1{1-p}}\cdot\DM\rfloor} 
\]
Then the interpretation of a Markov chain $C$ as a term is as follows, where final states are \emph{choices}.
\[
	\sem C~=~
	\left\{\begin{array}{ll}
		\term{s} & \text{if}~s~\in~F
	\\	{\colorterm\lfloor\D{{\sem{C(t_1)}}.p_1,..,{\sem{C(t_n)}}.p_n}\rfloor^s} & \text{where}~\delta(s)=\D{t_1.p_1,..,t_n.p_n}	
	\end{array}\right.
\]
That is, for a non-final state $s$ we inductively translate all children $t$ of $s$ as $\sem{C(t)}$, taking the chain $C(t)$ where $s$ now as a final state and $t$ the new initial state. The term for $C$ is then a term for the distribution $\delta(s)$, with the translation $\sem{C(t)}$ for each child $t$, looping on $s$.

\paragraph{Example}
Consider the following Markov chain $C$, where each arrow has probability $\frac13$.
\[
\vcenter{\hbox{
\begin{tikzpicture}[inner sep=2pt,outer sep=0pt,thick]
	\node[circle,draw] (a) at ( 90:1) {$a$} ;
	\node[circle,draw] (b) at (210:1) {$b$} ;
	\node[circle,draw] (c) at (330:1) {$c$} ;
	\node (i) at ( 90:2) {$i$} ;
	\node (j) at (210:2) {$j$} ;
	\node (k) at (330:2) {$k$} ;
	\begin{scope}[->,>=stealth]
	\draw (a.240) to[bend right] (b.60) ;
	\draw (b.0)   to[bend right] (c.180) ;
	\draw (c.120) to[bend right] (a.300) ;
	\draw (a.330) to[bend left]  (c.90) ;
	\draw (b.90)  to[bend left]  (a.210) ;
	\draw (c.210) to[bend left]  (b.330) ;
	\draw (a)--(i);
	\draw (b)--(j);
	\draw (c)--(k);
	\draw (120:2)--(a);
	\end{scope}
\end{tikzpicture}}}
\qquad
\begin{array}{l@{}l@{}l}
	S &~=~& \{a,b,c,i,j,k\}
\\[3pt]	s &~=~& a
\\[3pt]	F &~=~& \{i,j,k\}
\\[3pt]	\delta &~=~&  a \mapsto \D{~b./13~,~c./13~,~i./13~}
\\[3pt]            && b \mapsto \D{~a./13~,~c./13~,~j./13~}
\\[3pt]            && c \mapsto \D{~a./13~,~b./13~,~k./13~}
\end{array}
\]
Its interpretation is as follows, writing $\term{M+N+P}$ for the term $\colorterm\lfloor\D{M./13,N./13,P./13}\rfloor$.
\[
\begin{aligned}
	\sem C &~=~ \term{({\sem{C(b)}}+{\sem{C(c)}}+i)^a}
\\	       &~=~ \term{((a+{\sem{C(b)(c)}}+j)^b + (a+{\sem{C(c)(b)}}+k)^c+i)^a}
\\         &~=~ \term{((a+(a+b+k)^c+j)^b + (a+(a+c+j)^b+k)^c+i)^a}
\end{aligned}
\]
This term is typed as follows.
\[
\begin{aligned}
	    \term{(a+b+k)^c}         &: \type{\e => \e./13a + \e./13b + \e./13k}
\\[5pt]	\term{ a+(a+b+k)^c+j}~~~ &: \type{\e => \e./49a + \e./19b + \e./19k + \e./13j}
\\[5pt] \term{(a+(a+b+k)^c+j)^b} &: \type{\e => \e./12a + \e./18k + \e./38j}
\\[5pt] \term{(a+(a+c+j)^b+k)^c} &: \type{\e => \e./12a + \e./18j + \e./38k}
\\[5pt] \term{ (a+(a+b+k)^c+j)^b + (a+(a+c+j)^b+k)^c+i}~~~ &: \type{\e => \e./13a + \e./16k + \e./16j + \e./13i}
\\[5pt] \term{((a+(a+b+k)^c+j)^b + (a+(a+c+j)^b+k)^c+i)^a} &: \type{\e => \e./14k + \e./14j + \e./12i}
\end{aligned}
\]
The interpretation may be illustrated as follows, where the solid edges represent the tree-shape of the probabilistic sums, while the dashed edges represent the loops.
\[
\begin{tikzpicture}[x=24pt,y=-24pt,inner sep=2pt,outer sep=0pt,thick]
	\node[circle,draw] (a) at (1,0) {$a$};
	\node[circle,draw] (b) at (0,1) {$b$};
	\node[circle,draw] (c) at (2,1) {$c$};
	\node[circle,draw] (d) at (0,2.4) {$c$};
	\node[circle,draw] (e) at (2,2.4) {$b$};
	\node (i) at (1,-1)   {$i$};
	\node (j) at (-1,1)   {$j$};
	\node (k) at ( 3,1)   {$k$};
	\node (l) at (-1,2.4) {$k$};
	\node (m) at ( 3,2.4) {$j$};
	\begin{scope}[->,>=stealth]
	\draw (0,-1)--(a);
	\draw (a) to[bend right] (b);
	\draw (a) to[bend left]  (c);
	\draw (a) -- (i);
	\draw (b) -- (d);
	\draw (c) -- (e);
	\draw (b) -- (j);
	\draw (c) -- (k);
	\draw (d) -- (l);
	\draw (e) -- (m);
	\begin{scope}[dash pattern=on 3pt off 2pt]
	\draw (b.120) to[bend left]  (a.150);
	\draw (c.60)  to[bend right] (a.30);
	\draw (d.120) to[bend left]  (b.240);
	\draw (e.60)  to[bend right] (c.300);
	\draw (d.60)  to[bend left=10]  (a.240);
	\draw (e.120) to[bend right=10] (a.300);
	\end{scope}
	\end{scope}
\end{tikzpicture}
\]
As the illustration suggests, the formalism may express any probabilistic tree with back-pointers. It may also express directed acyclic graphs, which yields a smaller interpretation of the Markov chain above, though still exponential, in the following term.
\[
	\term{ (b+c+i ; b -> a+c+j ; c -> (a+b+k ; b -> a+c+j)^c )^a }
\]
The types work out the same way.
\[
\begin{aligned}
	    \term{ b+c+i }~~~             &: \type{\e => \e./13b + \e./13c + \e./13i}
\\[5pt] \term{ b+c+i ; b -> a+c+j}~~~ &: \type{\e => \e./19a + \e./49c + \e./13i + \e./19j}
\\[5pt] \term{(a+b+k ; b -> a+c+j)^c} &: \type{\e => \e./12a + \e./18j + \e./38k}
\\[5pt] \term{ b+c+i ; b -> a+c+j ; c -> (a+b+k ; b -> a+c+j)^c}     &: \type{\e => \e./13a + \e./13i + \e./16j + \e./16k}
\\[5pt] \term{(b+c+i ; b -> a+c+j ; c -> (a+b+k ; b -> a+c+j)^c )^a} &: \type{\e => \e./12i + \e./14j + \e./14k}
\end{aligned}
\]

%----------------------------------------------------------------------------------------------------
\newpage
\subsection{AST vs PAST}

The following term $\term Z$ generates a geometric series of Church numerals.
\[
	\term{Z}~=~\term{ <f>.<x>.[x].~(<n>.~([n].*~+~[[n].f].i))^i ; <m>.m }~\sim~\D{\cno./12,~\cni./14,~\cnii./18,~\dots}
\]
Let $\term M=\term{<n>.([n].*~+~[[n].f].i)}$ so that the above term is $\term{ <f>.<x>.[x].M^i ; <m>.m }$. For a given $\term N$, the term$\term{[N].M^i}$ evaluates as follows.
\[
\begin{aligned}
\term{[N].M^i}  
 ~ \rws ~ & \term{[N].M;i->M^i}
\\    = ~ & \term{[N].<n>.([n].*~+~[[n].f].i);i->M^i}
\\  \rw ~ & \term{([N].*~+~[[N].f].i);i->M^i}
\\ \sim ~ & \term{[N].* ~+~ [[N].f].M^i}
\end{aligned}
\]
The overall term $\term{ <f>.M^i ; <x>.x }$ then generates the geometric series as follows.
\[
\begin{array}{@{}r@{}l@{}}
	       & \term{ <f>.<x>.[x].M^i ; <m>.m }
\\  ~\rws~ & \term{ <f>.<x>.([x].* ; [[x].f].M^i) ; <m>.m }
\\ 	~\sim~ & \term{ <f>.<x>.[x].<m>.m ~+~ (<f>.<x>.[[x].f].M^i ; <m>;m) }
\\  ~\rw ~ & \term{ \cno ~+~ (<f>.<x>.[[x].f].M^i ; <m>;m) }
\\  ~\rws~ & \term{ \cno ~+~ (\cni ~+~ (<f>.<x>.[[[x].f].f].M^i ; <m>;m)) }
\end{array}
\]
Church numerals allow exponentiation by $\underline{m^n}=\cnn\,\cnm$. The term $\term{[\cnii].Z}$ then generates the following series.
\[
	\term{[\cnii].Z}~\sim~\D{\cni./12~,~\cnii./14~,~\cniv./18~,~..}
\]
Since Church numerals are in unary notation, the terms in the series grow exponentially. Stochastically generating a Church numeral with $\term{[\cnii].Z}$ is thus almost-surely terminating, but not \emph{positively} so.

There are two minor caveats. First, the above computations use reduction $\rw$ and equivalence $\sim$, but not evaluation on the machine. Second, since the calculus encodes constants as choices, it only admits finite datatypes. One may address both issues by introducing a natural numbers type $\type\N$ with primitive operations $\term{`+:\N\,\N=>\N}$ etc. Then the following term evaluates on the machine to return a single natural number according to the above series. (This has been confirmed experimentally.)
\[
	\term{[[0].*].[<n>.n ; [1].`+].[\cnii].Z}
\]
Note that Church numerals $\term{\cnn}$ follow a call--by--name semantics, while primitive operations on natural numbers have a call--by--value semantics. This is reflected in the base value $\term{[0].*}$ and successor function $\term{<n>.n ; [1].`+}$, which incorporate the encoding of call--by--value into call--by--name.

It remains to show that the above terms can be typed. First, for any type $\type{s=>t}$, and indeed any type $\type{!s=>!t_I}$, the Church numerals and the term $\term Z$ can be typed as follows.
\[
\begin{aligned}
	    \term{ n }                &: \type{s=>t}
\\[5pt] \term{ f }                &: \type{(s=>t)\,s => t}
\\[5pt] \term{ [[n].f].i }        &: \type{\e => (s=>t).i}
\\[5pt] \term{ [n].*~+~[[n].f].i} &: \type{\e => (s=>t)./12* + (s=>t)./12i}
\\[5pt] \term{M}~=~\term{<n>.([n].*~+~[[n].f].i)} &: \type{(s=>t) => (s=>t)./12* + (s=>t)./12i}
\\[5pt] \term{M^i}                &: \type{(s=>t) => (s=>t)}
\\[5pt] \term{Z}~=~\term{ <f>.<x>.[x].~M^i ; <m>.m } &: \type{((s=>t)\,s => t)~(s=>t)~s~=>~t}
\end{aligned}
\]
It follows that $\term{[\cnii].Z}$ can be typed with the same type. The types of the natural number constructions are as follows.
\[
	\term{[0].* : \e=>\N} \qquad
\qquad
	\term{<n>.n ; [1]. `+ : (\e=>\N)=>\N}
\]
This gives the following overall type. 
\[
	\term{[[0].*].[<n>.n ; [1].`+].[\cnii].Z : \e=>\N}
\]
The Church numeral types $\type{((s=>t)\,s => t)~(s=>t)~s=>t}$ of $\term{\cnii}$ and $\term{Z}$ here are specialised as follows: that for $\term{\cnii}$ by replacing $\type{s=>t}$ with $\type{\e=>\N}$, and that for $\term Z$ by replacing $\type{s=>t}$ with $\type{(\e=>\N)=>\N}$.
\[
\begin{aligned}
	    \term{\cnii} &: \type{((\e=>\N)=>\N)\,(\e=>\N)=>\N}
\\[5pt]	\term{Z}     &: \type{(((\e=>\N)=>\N)\,(\e=>\N)=>\N)~((\e=>\N)=>\N)~(\e=>\N)~=>~\N}
\end{aligned}
\]
The existence of this typed term gives the following.

\begin{proposition}
The probabilistically typed FMC expresses terms that are not positively almost-sure terminating.
\end{proposition}

%----------------------------------------------------------------------------------------------------

\end{document}

%====================================================================================================

% Bounded evaluation
%
\begin{definition}
The \emph{bounded evaluation} relation $\eval[m]SM\DT$ is defined inductively by the following rules.
\[
\begin{array}{@{}c@{\qquad}c@{\qquad}c@{}}
	\infer{\eval[m]Si{\D{S.1i}}}{}
&	\infer{\eval[m]{S\,\term N}{<x>.M}\DT}{\eval[m]S{\{N/x\}M}\DT}
&	\infer{\eval[m]R{M;i->N}{\DS + \sum_{k\leq n}p_k\cdot\DT_k}}{
    	\eval[m]RM {\DS + \D{S_k.p_ki}_{k\leq n}}
    	&& 
    	\{\eval[m]{S_k}N{\DT_k}\}_{k\leq n}
    }
\\ \\
	\infer{\eval[m]S{M}\varnothing}{}
&	\infer{\eval[m]S{[N].M}\DT}{\eval[m]{S\,\term N}M\DT}
&	\infer{\eval[m]S{?a.M}{\frac12\DT_\1+\frac12\DT_\0}}{\eval[m]S{\{\1/a\}M}{\DT_\1} && \eval[m]S{\{\0/a\}M}{\DT_\0}}
\end{array}	
\]
\[
\infer{\eval[0]S{M^i}\varnothing}{}
\qquad
\infer{\eval[m+1]R{M^i}{\DS + \sum_{k\leq n}p_k\cdot\DT_k}}{
    	\eval[m+1]RM {\DS + \D{S_k.p_ki}_{k\leq n}}
    	&& 
    	\{\eval[m]{S_k}{M^i}{\DT_k}\}_{k\leq n}
    }
\]
\end{definition}

%\[
%	\D{S_k.p_ki_k}_{k\leq n}\,\Downarrow\,\term{i->M}\,=\,\sum_{k\leq n}[p_k\cdot(S_k\Downarrow\term M)\mid i_k=i]+[S_k\cdot p_ki_k\mid i_k\neq i]
%\]

We will consider the system without the rule for $\eval[m]S{M}\varnothing$, which simplifies the system in two ways: evaluation is a function, and return types and return values are distributions.

\begin{proposition}
Evaluation $\eval[m]SM\DT$ is a function.
\end{proposition}

By the above proposition, we may write $\eval[m]SM{}$ to mean $\DT$.

\begin{proposition}
$(\eval[m]SM{}) \subseteq (\eval[m+1]SM)$
\end{proposition}

\begin{definition}
\emph{Evaluation} is defined as the limit of bounded evaluation:
\[
	\eval SM{}~=~\sup_{m\in\mathbb N}(\eval[m]SM{})
\]
\end{definition}

\end{document}

over triples of a trace $t$, a stack $S$, and a choice $i$, with the additional conditions that 1) if $n\neq m$ then $t_n\neq t_m$, and 2) $p_n=p(t_n)$ where $p(t)=(\frac12)^{|t|}$ is the \emph{probability} of a trace $t$, with $|t|$ its length. That is, traces uniquely identify elements of the distribution and determine their probability. Since probabilities are determined by traces, we may omit them and write $(t,S,i)$ for $(t,S,i)\cdot p$.

We write $s\cdot\DS$ for the distribution $\DS$ where every trace is prefixed by $s$:
\[
	s\cdot\Ds{(t_n,S_n,i_n)}{n\in\N}~=~\Ds{(s\,t_n,S_n,i_n)}{n\in\N}
\]
%The \emph{traces} $\floor{\DS}_t$ of a distribution $\DS$ are the traces in its support.
%\[
%	\floor{\DS}_t~=~\{ t \mid (t,S,i)\in\floor{\DS}\,\}
%\]
Finally, we define the \emph{restriction} $\DS|i$ of a distribution to a choice $i$, and its \emph{exclusion} $\DS\less i$ of $i$, as follows.
\[
	\DS|i~=~\Ds{ (t,S,j)\in\DS \mid j=i }{} \qquad 	\DS\less i~=~\Ds{ (t,S,j)\in\DS \mid j\neq i }{}
\]

Let the distribution $\DR SM$ be that over the runs for $\term M$ with stack $S$, as follows.
\[
	\DR SM~=~\D{ (t,T,i) \mid {\step= \e SM\e tTi\e}~}
\]

\end{document}